---
title: "Class 5: Data Visualization with ggplot"
author: "Barry"
format: pdf
toc: true
---

# Our first plot

R has base graphics

```{r}
head(cars)
```

```{r}
plot(cars)
```

How would I plot this with `ggplot2`?
NO!

We need to install and load the ggplot2 package first.
To install any package in R we use the `install.packages()` function.

```{r}
# install.packages("ggplot2")
```

Before I can use this package I need to load it with a `library()` call.

```{r}
library(ggplot2)
ggplot(cars)
```

Every ggplot needs at least 3 layers:

- **Data** (i.e. the data.frame we have),
- **Aes** (the aesthetic mapping of our data to what we want to plot)
- **Geoms** (How we want to plot this stuff!)

```{r}
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point()
```


Add annother geom

```{r}
ggplot(data=cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

## A more intresting plot

First read data from online

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

Q1. How many genes are in this dataset

```{r}
nrow(genes)
```

There are `r nrow(genes)` genes in this data set.


```{r}
p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State, names=Gene) +
  geom_point()

p
```


# Using plotly for interactive plots

```{r}
library(plotly)


ggplotly(p)
```
## Try a different plot
```{r}
p2 <- ggplot(data=mpg) +
  aes(x=displ, y=hwy, color=class,
      shape=factor(cyl), names=model) + 
  geom_point()

p2
```
and plotly version

```{r}
ggplotly(p2)
```

